---
title: "Variation in the salience of cultural issues
over time and across countries"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE,
                      fig.showtext=FALSE,
                      fig.align='center')
knitr::opts_knit$set(root.dir = "C:/Users/voyta/Desktop/Thesis")

```

```{r}
library(tidyverse)
library(plm)
library(gplots)
library(tseries)
library(huxtable)
library(ggpubr)
library(ggeffects)
library(jtools)
library(ggrepel)
library(fixest)

```

```{r data_prep, cache=TRUE, include=FALSE}

df_party <- read.csv("data/df_party_coded.csv") # party level data

df_elect <- read.csv("data/df_elect_cl.csv") # election level data


#long dataset for graphing
df_long <- df_elect %>% 
  pivot_longer(cols = uncod_a:non_econ_all_a,
               names_to = "issue_area",
               values_to = "share")

# category names for later 
df_long <- df_long  %>% 
 mutate(category = fct_recode(issue_area,
                              "Aggregate economic" = "econ_agg_a",
                              "Nationalism, multiculturalism,immigration, and minorities + morality issues" = "nat_multi_mor_a", 
                              
                            "Aggregate cultural" =  "cult_agg_a",
                            "All non-economic" = "non_econ_all_a",
                            "Morality" = "mor_a",
                            "Nationalism, multiculturalism, immigration, and minorities" = "nat_multi_a",
                            "Environment" = "envi_a",
                            "Welfare and labour" = "welf_a",
                            "Economy" = "econ_a", 
                            "EU" = "EU_a"))

df_elect <- df_elect %>% 
  group_by(countryname) %>%
  mutate(election_number = row_number()) %>%
  ungroup()
  
                            
                      
```

## [Salience of cultural vs economic issues]{.underline}

**OI: At the aggregate level of party systems, cultural issues have been relatively less salient than economic issues throughout the whole studied period.**

##### Comparison of the salience of cultural issues in the aggregate (including multiculturalism, morality, EU, environment, law and order etc.) and aggregate economic issues at the level of elections:

```{r}
df_long %>% 
  filter(decade != "2020s") %>%
  filter(issue_area %in% c("econ_agg_a", "cult_agg_a")) %>%
  ggplot(aes(x = category, y = share)) + 
  geom_boxplot() +
  geom_point(position = position_jitter(),
             alpha = 0.3) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  facet_wrap(~decade) +
  theme_pubclean() +
  labs( x= "Issue area", y = "Mean salience (%)",
        title = "Salience of different issue areas at elections")
  
```

## [Is the salience of cultural issues increasing over time?]{.underline}

**OI: The relative salience of cultural issues is increasing over time in CEE party systems.**

##### There are no strong trends in the data:

##### Line graph with loess

```{r line_graph, echo=FALSE}


cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#CC79A7")


df_long %>% 
  filter(issue_area %in% c("welf_a", "econ_a", "nat_multi_a", "EU_a", "envi_a","mor_a")) %>%
  ggplot(aes(x=year, y=share, color = category))+
  geom_smooth(aes(linetype=category),method = "loess", se=T, span =0.8)+
  geom_point(aes(shape = category), alpha = 0.3) +
  scale_color_manual(values = cbbPalette)+
  scale_x_continuous(breaks = seq(1990,2020, by = 5))+
  scale_y_continuous(breaks = seq(0,55, by=5))+
  theme_classic2()+ 
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5),
        legend.text = element_text(size = 8)) +
  labs(title = "Development of issue content over time
       (election level)", x= "Year", y= "Mean salience (%)")
```

##### OLS of party issue salience on year (country dummies included, not shown):

```{r}
m1_y <- lm(nat_multi ~ year + factor(country),
         data = df_party)

m2_y <-  lm(mor ~ year + factor(country),
         data = df_party)

m3_y <-  lm(envi ~ year + factor(country),
         data = df_party)

m4_y <-  lm(EU ~ year + factor(country),
         data = df_party)

export_summs(m1_y, m2_y, m3_y, m4_y, robust = T,
             model.names = c("Multiculturalism",
                        "Morality",
                        "Environment",
                        "EU"), 
             coefs = "year")
```

##### Predictions from the models above:

```{r}
pred1 <- ggpredict(m1_y, terms = "year")
pred2 <- ggpredict(m2_y, terms = "year")
pred3 <- ggpredict(m3_y, terms = "year")
pred4 <- ggpredict(m4_y, terms = "year")


pred1$model <- "Multiculturalism"
pred2$model <- "Morality"
pred3$model <- "Environment"
pred4$model <- "EU"


all_preds <- bind_rows(pred1, pred2, pred3, pred4)

# Create faceted plot
ggplot(all_preds, aes(x = x, y = predicted)) +
  geom_line() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  facet_wrap(~ model, scales = "free_y") +
  labs(x = "Year", y = "% of quasisentences") +
  theme_classic2()
```

## [Was the migration crisis associated with an increase of multiculturalism salience?]{.underline}

**OI: The period following the migration crisis associated with an increase in the salience of multiculturalism and immigration.**

##### Salience of multiculturalism at elections over time:

```{r message=FALSE, warning=FALSE}


points <- df_elect %>% 
  filter(nat_multi_a >= 6) 

df_elect %>% 
  ggplot(aes(x=year, y=nat_multi_a))+
  geom_point(shape = 1) +
  geom_text_repel(data = points,
            aes(label = elect_label),
            alpha= 0.5,
            size = 3) +
  geom_vline(xintercept = 2015, 
             linetype = "dashed") +
  geom_smooth(linetype = "dashed",
              alpha = 0.8,
              se = F,
              color = "grey") +
  annotate("text", x=2015, y=20, label="2015", angle=0) +
  scale_y_continuous(breaks = seq(0,30, by=5))+
  scale_x_continuous(breaks = seq(2000, 2020, by =2),
                     limits = c(2000,2020)) +
  theme_classic2() + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5)) +
  labs(x= "Year", y= "Mean salience (%)")

```

##### Salience of multiculturalism - by country

```{r}

points <- df_elect %>% 
  filter(nat_multi_a >= 6) 

df_elect %>% 
  ggplot(aes(x=year, y=nat_multi_a))+
  geom_point(shape = 1) +
  geom_line() +
  geom_vline(xintercept = 2015, 
             linetype = "dashed") +
  facet_wrap(~countryname) +
  annotate("text", x=2015, y=20, label="2015", angle=0) +
  scale_y_continuous(breaks = seq(0,40, by=5))+
  scale_x_continuous(breaks = seq(2000, 2020, by =4),
                     limits = c(2000,2020)) +
  theme_classic2() + 
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5)) +
  labs(x= "Year", y= "Mean salience (%)")

```

## [Does the salience of cultural issues differ conditional on ethnic minorities, religion, or communist legacies?]{.underline}

**OI: The salience of cultural issues should be higher in countries that have politically significant ethnic minorities**

**BUT also in those that have a legacy of national accomodative communism even if ethnic minorities are absent**

##### Country-level plots:

```{r}

df_long %>% 
  filter(issue_area %in% c("nat_multi_a", "mor_a")) %>%
  ggplot(aes(x=year, y=share, color = category))+
  geom_smooth(aes(linetype=category),method = "loess", se=T, span =0.8)+
  geom_point(aes(shape = category), alpha = 0.3) +
  scale_color_manual(values = cbbPalette)+
  scale_y_continuous(breaks = seq(0,30, by=5))+
  theme_classic2()+ 
  theme(legend.position = "bottom",
         axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~countryname)+
  labs(title = "Development of issue content in CEE
       morality and multiculturalism",
    x= "Year", y= "Mean salience (%)")


```

```{r eval=FALSE, include=FALSE}
df_elect %>%
  ggplot(aes(x=year, y= cult_agg_a))+
  geom_smooth(method = "loess", se=T, span =0.8)+
  geom_point(shape = 1, alpha = 0.3) +
  scale_y_continuous(breaks = seq(0,30, by=5))+
  theme_classic2()+ 
  theme(legend.position = "bottom",
         axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust = 0.5)) +
  facet_wrap(~countryname)+
  labs(title = "Development of issue content in CEE - Aggregate cultural issues", x= "Year", y= "Mean salience (%)")
```

##### Differences in the mean salience of multiculturalism at elections by countries with different communist legacies and ethnic structure:

```{r echo=FALSE}


df_elect %>%
  ggplot(aes(x = as.factor(minority_legacy), y = nat_multi_a)) +
  geom_boxplot() + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  geom_point(position = position_jitter(),
             alpha = 0.3,
             shape = 1) +
  geom_text_repel(aes(label = elect_label),
            alpha = 0.3)+
  labs(y = "Mean salience (%)", 
       x = "ethnic minority presence and legacy")+
  theme_pubclean()

```

```{r eval=FALSE, include=FALSE}

# Party-level differences in multiculturalism
df_party %>%
  ggplot(aes(x = as.factor(minority_legacy), y = nat_multi)) +
  geom_boxplot() + 
  geom_point(position = position_jitter(),
             alpha = 0.3,
             shape = 1) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_y_continuous(breaks = seq(0,60, by = 10))+
  labs(y = "% of quasisentences", 
       x = "ethnic minority presence and legacy")+
  theme_pubclean()

```

##### [Religion and the salience of morality issues]{.underline}

**OI: Countries' religiosity levels are positively associated with the salience of cultural issues, especially those related to morality.**

There is no significant correlation between countries' religiosity levels and the salience of morality issues at elections:

```{r eval=FALSE, include=FALSE}
cor.test(df_elect$share_rel, df_elect$mor_a)

```

```{r}
df_elect %>% 
  ggplot(aes(x = share_rel, y = mor_a)) +
  geom_point() +
  geom_text_repel(aes(label = elect_label),
                  alpha = 0.3,
                  size = 2)+
  geom_smooth(method = "lm",
              color = "steelblue") +
  theme_pubclean() +
  labs(y = "Mean salience (%)",
       x = "Share of the population that is religious
                (European Social Survey data)",
       title = "Religiosity and the salience of morality issues at elections")
```

Note: Data on issue salience is from the Manifesto project dataset:

Lehmann, Pola, Simon Franzmann, Denise Al-Gaddooa, Tobias Burst, Christoph Ivanusch, Sven Regel, Felicia Riethmüller, et al. 2024. "Manifesto Project Dataset." Manifesto Project. <https://doi.org/10.25522/MANIFESTO.MPDS.2024A.>

Data on religiosity is from the European Social Survey Wave 11 integrated file:

European Social Survey European Research Infrastructure (ESS ERIC). (2024). ESS11 integrated file, edition 2.0 [Data set]. Sikt - Norwegian Agency for Shared Services in Education and Research. <https://doi.org/10.21338/ess11e02_0>

Communist legacies are coded based on:

Kitschelt, Herbert, Zdenka Mansfeldova, Radoslaw Markowski, and Gabor Toka. 1999. *Post-Communist Party Systems: Competition, Representation, and Inter-Party Cooperation*. Cambridge Studies in Comparative Politics. Cambridge: Cambridge University Press. <https://doi.org/10.1017/CBO9781139175173>.

Countries with politically significant ethnic minorities are coded based on:

Rovny, Jan. 2024. *Ethnic Minorities, Political Competition, and Democracy: Circumstantial Liberals*. Transformations in Governance. Oxford: Oxford University Press.
